|

### 2021-12-10 05-Chp-CPRD-Analysis.Rmd
End of day:

@@

Finally, as with many studies of dementia, there is a risk of reverse causation in my analysis. Dementia and associated conditions have a long prodromal period, during which preclinical disease could cause indications for the prescription of a lipid-regulating agent.

>>>>

<!-- TODO Talk here about the reduced biological validity of  -->

@@

Finally, as with many studies of dementia, there is a risk of reverse causation in my analysis. Dementia and associated conditions have a long prodromal period, during which preclinical disease could cause indications for the prescription of a lipid-regulating agent. Enforcing a minimum period of follow-up would address this limitation, but the use of a time-varying treatment indicator in this analysis prevents the use of this approach (see Appendix \@ref(cprd-min-fu) for a fuller discussion of this topic).

### 2021-12-10 06-Chp-IPD-Analysis.Rmd
End of day:

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed (see Appendix \@ref(ipd-email-collab) for a copy of the correspondence sent to potential collaborators) in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

A range of additional variables were included in the analysis, intended to address the potential for confounding. With an awareness that discrepancies in available covariates is a common issue in IPD studies, I defined an idealised set of covariate domains to be age, sex, education, BMI, _Apo_$\mathcal{E}4$ status, smoking/alcohol status, ethnicity, and prevalent cardiovascular disease. This set covers key risk factors for dementia/Alzheimer's disease in addition to general cardiovascular risk factors. Details on how these variables were coded, given the available data, are presented in Section \@ref(ipd-covar-definition).

@@

<!-- TODO Need average follow-up by cohort here - or rather put it in the table -->
<!-- TODO Need total number of Azd cases in whitehall in the paragraph above-->

@@

```{r covariateSummary-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

>>>>

The first approach focused on previously analysed observational prospective cohort studies examining the effect of blood lipid levels on dementia outcomes, as identified by the systematic review presented in Chapters \@ref(sys-rev-methods-heading) & \@ref(sys-rev-results-heading). The data sources used in each of these cohort analyses were screened against the criteria listed in the previous section, and eligible cohorts were approached for data access. In the first instance, the first/corresponding author of the publication was emailed <!-- TODO (see Appendix \@ref(ipd-email-collab) for a copy of the correspondence sent to potential collaborators) --> in Autumn 2020. If this approach did not elicit a response within two months, the last author was contacted, on the basis that the first/corresponding author may have been a more junior member of the research group who had moved to a different institution.

@@

A range of additional variables were included in the analysis, intended to address the potential for confounding. With an awareness that discrepancies in available covariates is a common issue in IPD studies, I defined an idealised set of covariate domains to be age, sex, education, BMI, _Apo_$\mathcal{E}4$ status, smoking/alcohol status, ethnicity, and prevalent diabetes or cardiovascular disease. This set covers key risk factors for dementia/Alzheimer's disease in addition to general cardiovascular risk factors. Details on how these variables were coded, given the available data, are presented in Section \@ref(ipd-covar-definition).

<!-- TODO Need to go through and make sure that I have updated the covariate list appropriately -->

@@

@@

```{r covariateSummary-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

<!-- TODO Note here about how the IHD variable in CAPS has a note saying there might be an issue with the variables, while the Whitehall study uses narrow range of codes to define CHD. EPIC is selfreported angina/mi/arrythmia -->

@@

<!-- TODO If using fixed effect model, need to udate subgroup labels to reflect this -->

### 2021-12-10 front-and-back-matter/99-appendices.Rmd
End of day:

*	*Page, M.J., Moher, D., Bossuyt, P., Boutron, I., Hoffmann, T., Mulrow, C., Shamseer, L., Tetzlaff, J., Akl, E., Brennan, S.E., Chou, R., Glanville, J., Grimshaw, J., Hróbjartsson, A., Lalu, M.M., Li, T., Loder, E., Mayo-Wilson, E., McDonald, S., __McGuinness, L.A.__, Stewart, L., Thomas, J., Tricco, A., Welch, V.A., Whiting, P., McKenzie, J. (2021). PRISMA 2020 statement: updated guidelines for reporting systematic reviews and meta-analyses. BMJ 372:n71 [10.1136/bmj.n71](https://doi.org/10.1136/bmj.n71)*

@@

*	*Schmidt, L., Olorisade, B.K., __McGuinness, L.A.__, Thomas, J. and Higgins, J.P. (2020). Data extraction methods for systematic review (semi) automation. F1000Research, 10:401. DOI: &nbsp; [10.12688/f1000research.51117.1](https://doi.org/10.12688/f1000research.51117.1)*

@@

@@

@@

```{r mrTool-table, message=FALSE, results="asis", echo = FALSE}

mrTool_table <- read.csv(here::here("data/sys-rev/mrRobTool.csv")) %>%
  select(1:5) %>%
  rename("Bias domain" = Bias.domain)

if(doc_type == "docx") {
  knitr::kable(mrTool_table, caption = "(ref:mrTool-caption)")
} else{
  knitr::kable(
    mrTool_table,
    format = "latex",
    caption = "(ref:mrTool-caption)",
    caption.short = "(ref:mrTool-scaption)",
    longtable = TRUE,
    booktabs = TRUE
  ) %>%
    row_spec(0, bold = TRUE) %>%
    column_spec(column = 1:5, width = paste0(32/5,"em")) %>%
    row_spec(2:nrow(mrTool_table)-1, hline_after = TRUE) %>%
    kable_styling(latex_options = c("HOLD_position","repeat_header"))
}

@@

@@

### Email and documents sent to potental collaborators {#ipd-email-collab}

__To be copied from email__

@@

@@

<!----------------------------------------------------------------------------->
(ref:ipdDementiaDef-caption) Definition of dementia outcomes across cohorts included in the individual participant data meta-analysis

@@

In an attempt to inform the prior distributions used in the triangulation exercise presented in Chapter \@ref(tri-heading), I first averaged the mean and variance values across the four experts for each source of bias/indirectness within each study. I averaged the absolute value of the mean, as I wanted to obtain the average adjustment regardless of direction of adjustment which was indicated using the sign of the mean in the original analysis).

I then took the average and maximum of these averaged mean and variance values. The distributions arising from this analysis are presented in Table \@ref(tab:turnerEstimates-table). The two step approach (averaging across assessors before taking the average/maximum across sources of bias) was important. While the final mean values will be the same whichever approach is taken, in a one-step process the maximum value can be artificially inflated by a single expert's estimate. Averaging across experts for a given bias in a given study before taking the maximum value across biases and studies prevented this.

<!----------------------------------------------------------------------------->
(ref:turnerEstimates-caption) Mean and maximum distributions for bias and indirectness obtained from data resulting from the elicitation exercise reported in _Turner et al_[@turner2009]

@@

```{r turnerEstimates-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

As discussed in the main text, I defined two sets of additive bias distributions as a sensitivity analysis. The results for the two case studies under the two scenarios are presented in the figures below.

(ref:fpLdlAdComparison-cap) Comparison of bias-/indirectness-adjusted meta-analyses of LDL-c/Alzheimer's disease, using two different scenarios for the relationship of levels of bias. The left forest plot shows the result where the difference in mean value between the "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. In the right plot, the difference in mean value between "High" and "Moderate" is twice that between "Moderate" and "Low".

(ref:fpLdlAdComparison-scap) Comparison of bias-/indirectness-adjusted meta-analyses under two different scenarios of prior distributions of bias.

@@

(ref:fpTgVaDComparison-cap) Comparison of bias-/indirectness-adjusted meta-analyses of triglycerides/vascular dementia, using two different scenarios for the relationship of levels of bias. The left forest plot shows the result where the difference in mean value between the "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. In the right plot, the difference in mean value between "High" and "Moderate" is twice that between "Moderate" and "Low".

@@

All software projects in these thesis attempt to conform to minimal best practices for research computing. [@wilson2014; @wilson2017]

Reproducible and science has been a key theme running through this thesis, as reflected by the development of an open source tool to help search medRxiv and bioRxiv preprint metadata. In line with this, an open source copy of the code used to produce this thesis is available on GitHub, as is the code used to perform the analysis contained within it.

Unfortunately, given the. However, as discussed in other chapters of this thesis and elsewhere, sharing of analysis code is a useful step towards transparency when the underlying datasets are not readily available.

Commentary on the fact that the best you can do is replicate vs reproducible (due the closed nature of the data).

One is the ability to recreate the results given the same data and code, the other is the ability to recreate the results given the same code but a different dataset. IN theory it is possible to gain access the dataset given the information presented in Chapter \@(ref:cprd-analysis-heading). However, access is dependency on an ISAC application to the managing body of the CPRD.

@@

\blandscape
<!----------------------------------------------------------------------------->
(ref:biasData-caption) biasData

@@

<!----------------------------------------------------------------------------->
(ref:biasDirectionData-caption) biasDirectionData

@@

>>>>

*	*Page, M.J., Moher, D., Bossuyt, P., Boutron, I., Hoffmann, T., Mulrow, C., Shamseer, L., Tetzlaff, J., Akl, E., Brennan, S.E., Chou, R., Glanville, J., Grimshaw, J., Hróbjartsson, A., Lalu, M.M., Li, T., Loder, E., Mayo-Wilson, E., McDonald, S., __McGuinness, L.A.__, Stewart, L., Thomas, J., Tricco, A., Welch, V.A., Whiting, P., McKenzie, J. (2021). PRISMA 2020 statement: updated guidelines for reporting systematic reviews and meta-analyses. BMJ 372:n71. DOI:  [10.1136/bmj.n71](https://doi.org/10.1136/bmj.n71)*

@@

*	*Schmidt, L., Olorisade, B.K., __McGuinness, L.A.__, Thomas, J. and Higgins, J.P. (2020). Data extraction methods for systematic review (semi) automation. F1000Research, 10:401. DOI: [10.12688/f1000research.51117.1](https://doi.org/10.12688/f1000research.51117.1)*

@@

\blandscape

@@

A copy of the risk of bias tool used to assess Mendelian randomisation studies is presented in Table \@ref(mrTool-table), below.

@@

```{r mrTool-table, message=FALSE, results="asis", echo = FALSE}
```

\elandscape

@@

### Minimum follow-up period {#cprd-min-fu}

Enforcing a minimum period of follow-up is normally used to reduce the risk of reverse causation in an analysis by excluding events occurring within a certain time-frame post-exposure. With respect to dementia outcomes, which have a long prodomal period, enforcing a minimum period of follow-up can add biological validity to an analysis. Incident cases of dementia that occur within a few years of starting a drug may already be beyond effective preventive treatment at the time the drug was started, and so restricting the analysis to those with longer exposure to treatment may provide a more reliable estimate of the effect.

Panel A of Figure \@ref(fig:minFU) describes a standard analysis of an exposure with a minimum enforced follow-up period (in this case, 10 years). Participants enter the cohort and are classified at that point as either exposed or unexposed, for example, on the basis of a total cholesterol measurement. Participants are followed-up over time, but events occuring before the minimum period of follow-up are discarded.

<!----------------------------------------------------------------------------->
```{r minFollowUp, include = FALSE}
```

(ref:minFU-cap) __shortcap:__

(ref:minFU-scap) shortcap

```{r minFU, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:minFU-cap)', out.width='100%', fig.scap='(ref:minFU-scap)'}
knitr::include_graphics(file.path("figures/appendix/minFU.png"))
```
<!----------------------------------------------------------------------------->

However, applying a similar approach to the analysis performed in Chapter \@ref(cprd-analysis-heading), illustrated in Panel B of Figure \@ref(fig:minFU), is challenging. The use of a time updating treatment indicator results in two potential starting points for the 10-year lag: entry into the cohort or initiation of statin therapy. As the purpose of this analysis is to investigate the impact of statin treatment, starting the minimum period of follow-up from cohort entry is not helpful as it does not ensure that participants have been exposed to the drug for 10 years. Alternatively, defining the minimum period of follow-up from statin initiation cause issues around the treatment of time-at-risk before statin initiation and around the start point for the follow-up period in those never prescribed a statin. In summary, either approach is likely to be substantially biased.

With these caveats in mind, below I present a sensitivity analysis of the association of statins with dementia outcomes, enforcing a minimum 10-year period of follow-up from statin initiation in those prescribed a statin.

<!----------------------------------------------------------------------------->
(ref:cprdMinFu-cap) Estimates from the

(ref:cprdMinFu-scap) shortcap

```{r cprdMinFu, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:cprdMinFu-cap)', out.width='100%', fig.scap='(ref:cprdMinFu-scap)'}
knitr::include_graphics(file.path("figures/hold.png"))
```
<!----------------------------------------------------------------------------->

@@

<!-- ### Email sent to potental collaborators {#ipd-email-collab} -->

<!-- __To be copied from email__ -->

@@

<!----------------------------------------------------------------------------->
(ref:missingMatrix-caption) Overview of missing data across covariates included in IPD analysis. Note that LDL is missing completely for the Whitehall and CaPS datasets as this measure was derived from the measures for TC, HDL and TG following multiple imputation of missing values in these variables.

(ref:missingMatrix-scaption) missingMatrix

```{r missingMatrix-table, message=FALSE, results="asis", echo = FALSE}
```
<!----------------------------------------------------------------------------->

@@

<!----------------------------------------------------------------------------->
(ref:ipdDementiaDef-caption) Definition of dementia outcomes across cohorts included in the individual participant data meta-analysis. Note the codes used to define dementia and related outcomes in the HES data as part of the Whitehall II cohort were taken from a previously published analysis.[@russ2014]

@@

The resulting data for two domains of bias (attrition and detection) in one study are shown in Table \@ref(tab:addTable1-table).

&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable1-caption) Distributions of bias, defined by position and variance terms, assigned by four independent experts to attrition and detection bias in a single study. This data was obtained from a previous publication,[@turner2009] and represents an illustrative subset of the larger dataset.

(ref:addTable1-scaption) addTable1

```{r addTable1-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

In an attempt to inform the prior distributions used in the triangulation exercise presented in Chapter \@ref(tri-heading), I first averaged the position and variance values across the four experts for each source of bias/indirectness within each study. I took the average of the absolute of the position value, as I wanted to obtain the average adjustment regardless of direction. The results of this process for the example data presented below are shown in Table \@ref(tab:addTable2-table).

&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable2-caption) Position and variance values for two example bias domains (attritiona and detection), obtained by averaging the absolute of the values contained in \@ref(tab:addTable3-table) across the four reviewers.

(ref:addTable2-scaption) addTable1

```{r addTable2-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

I then took the average and maximum of the expert-averaged position and variance values. The results of this for the example data are presented in Table \@ref(tab:addTable3-table).

&nbsp;<!----------------------------------------------------------------------->  
(ref:addTable3-caption) Mean and maximum of position and variance values desribed in Table \@ref(tab:addTable3-table)

(ref:addTable3-scaption) addTable1

```{r addTable3-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

This two step approach (averaging across assessors before taking the average and maximum across domains/studies) was important. While the final average values will be the same whichever approach is taken, in a one-step process, the maximum value can be artificially inflated by a single expert's estimate. For example, in the illustrative data subset above, Assessor 2 assigns a distribution of N(-0.32,0.14), which would double the maximum position (0.16) and triple the maximum variance (0.0475) obtained when first averaging across experts.

The distributions arising from this applying this approach to the entire dataset, stratified by bias/indirectness and additive/proportional effects, are presented in Table \@ref(tab:turnerEstimates-table).

&nbsp;<!----------------------------------------------------------------------->  
(ref:turnerEstimates-caption) Mean and maximum distributions for bias and indirectness obtained from data resulting from the elicitation exercise reported in _Turner et al_[@turner2009]. Note I did not calculate values for additive indirectness, as in my analysis, all indirectness was assumed to be proportional.

@@

```{r turnerEstimates-table, message=FALSE, results="asis", echo = FALSE}
```
&nbsp;<!----------------------------------------------------------------------->

@@

As discussed in the main text, I defined two sets of additive bias distributions as a sensitivity analysis. The results for the two case studies under the two scenarios are presented in the figures below. Under the first, shown on the left of the subsequent  forest plot, the difference in the position value between the distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Under the second scenario, the difference in the position value between the distributions assigned to "High" and "Moderate" is twice that between "Moderate" and "Low".

(ref:fpLdlAdComparison-cap) Comparison of bias-/indirectness-adjusted meta-analyses of LDL-c/Alzheimer's disease, using two different scenarios for the relationship of levels of bias.

(ref:fpLdlAdComparison-scap) Comparison of bias-/indirectness-adjusted meta-analyses under two different scenarios of prior distributions of bias. Scenario 1: position value between the distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels

@@

(ref:fpTgVaDComparison-cap) Comparison of bias-/indirectness-adjusted meta-analyses of triglycerides/vascular dementia, using two different scenarios for the relationship of levels of bias. Scenario 1: position value between the distributions assigned to "High" and "Moderate" bias levels is the same as between the "Moderate" and "Low" levels. Scenario 2: position value between the distributions assigned to "High" and "Moderate" bias levels is twice that between "Moderate" and "Low" levels

@@

Reproducible and science has been a key theme running through this thesis, as reflected by the development of an open source tools to search preprinted literature, create visualisations of risk-of-bias data and to preform quantitative triangulation. In line with this, an open source copy of the code used to produce this thesis is available on GitHub, as is the code used to perform the analysis contained within it.

Unfortunately, given the nature of the data underpinning the CPRD (Chapter \@ref(cprd-analysis-heading)) and IPD (Chapter \@ref(ipd-heading)) analyses, it is not possible to make it publicly available. Access to the CPRD data is dependent on an application to its managing body, while acces to the cohorts within the DPUK requires approval from both the DPUK and inidividual cohort owners. IN the absence of data sharing, the best option to enable review and increase transparency is the open sharing of the code used to perform the analysis.

Finally, all software projects in these thesis attempt to conform to minimal best practices for research computing. [@wilson2014; @wilson2017]

@@

\blandscape
<!----------------------------------------------------------------------------->
(ref:biasData-caption) First five assessments contained in `data_rob2`, an example dataset for the Risk of Bias 2 tool for randomised controlled trials available via the `robvis` package.

@@

<!----------------------------------------------------------------------------->
(ref:biasDirectionData-caption) Example data illustrating the structure required for bias direction plots. $d_ij$, $d_id$, and $d_it$ define the extent (Serious, Moderate, Low), direction (right or left), and type (additive or proportional) of bias in the $i$th domain. Note that, because of the wide nature of the data, I have abbreviated the data in order to display it. A key for interpreting the data can be found below the table.

@@

### Simple visualisations

```{r robAppendixFigures, include = FALSE}
```

```{r, eval = FALSE}
rob_summary(data = data_rob2, tool = "ROB2")
```

<!----------------------------------------------------------------------------->
(ref:robSummary-cap) __shortcap:__

(ref:robSummary-scap) shortcap

```{r robSummary, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:robSummary-cap)', out.width='80%', fig.scap='(ref:robSummary-scap)'}
knitr::include_graphics(file.path("figures/appendix/robSummary.png"))
```
<!----------------------------------------------------------------------------->

```{r, eval = FALSE}
rob_traffic_light(data = data_rob2,
                  tool = "ROB2",
                  psize = 10)
```

<!----------------------------------------------------------------------------->
(ref:robTraffic-cap) __shortcap:__

(ref:robTraffic-scap) shortcap

```{r robTraffic, echo = FALSE, results="asis", fig.pos = "H", fig.cap='(ref:robTraffic-cap)', out.width='80%', fig.scap='(ref:robTraffic-scap)'}
knitr::include_graphics(file.path("figures/appendix/robTraffic.png"))
```
<!----------------------------------------------------------------------------->
